---
title: "2021_07_06_tidy_tuesday"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
theme_set(theme_light())
```

## Load data

```{r}
holidays <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-06/holidays.csv')

```

## Which country colonized others the most?

```{r}
holidays %>% 
    filter(!is.na(independence_from)) %>% 
    mutate(independence_from = case_when(
        str_detect(independence_from, ".*United Kingdom.*") ~ "United Kingdom",
        str_detect(independence_from, "Kingdom of Great Britain") ~ "United Kingdom",
        str_detect(independence_from, ".*France.*") ~ "France",
        str_detect(independence_from, ".*Spanish.* | Spain") ~ "Spain",
        str_detect(independence_from, ".*Soviet.*") ~ "Soviet Union",
        str_detect(independence_from, ".*Japan.*") ~ "Japan",
        TRUE ~ independence_from)) %>% 
    count(independence_from, sort = TRUE) %>% 
    print(n = Inf)
```

* United Kingdom

## How many countries declared independence from Qing China?

```{r}
holidays %>% 
    filter(str_detect(independence_from, "Qing.*"))
```

* Mongolia

## How many countries declared independence from US?

```{r}
holidays %>% 
    filter(str_detect(independence_from, "United States")) 
```

* Austria (US is one of the allies) and Micronesia


## Top 10 countries with the earliest independence day

```{r}

holidays %>% 
    slice_min(date_parsed, n = 10) %>% 
    relocate(country, date_parsed, independence_from) %>% 
    ggplot(aes(date_parsed, country)) +
    geom_point() +
    geom_text(aes(label = date_parsed), size = 2.5,
              vjust = 1.5) +
    facet_wrap(~independence_from) +
    labs(x = "Date of Independence",
         y = "",
         title = "Top 10 countries with the earliest independence day",
         subtitle = "Faceted by where they get independence from")
```

* Switzerland was the first country to get independence in August 1, 1291. Among the countries who got independence early, most of them got independence from Spanish Empire.

## Top 10 countries with the latest independence day

```{r}

holidays %>% 
    slice_max(date_parsed, n = 10) %>% 
    relocate(country, date_parsed, independence_from) %>% 
    ggplot(aes(date_parsed, country)) +
    geom_point() +
    geom_text(aes(label = date_parsed), size = 2.5,
              vjust = 1.5) +
    facet_wrap(~independence_from) +
    labs(x = "Date of Independence",
         y = "",
         title = "Top 10 countries with the latest independence day",
         subtitle = "Faceted by where they get independence from")
```

* South Sudan was the last country to get independence in July 9, 2011. Among the countries who got independence late, most of them got independence from Soviet Union.


## The most common holiday name for independence 

```{r}
holidays %>% 
    count(name_of_holiday, sort = TRUE)
```

* Independence day.
