---
title: "TidyTemplate"
date: 2021-02-23
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(tidytuesdayR)
theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-02-23")
earn <- tt$earn
employed <- tt$employed
```


# EDA on employed

```{r}
employed %>% 
    skimr::skim()

employed_df <- employed %>% 
    mutate_if(is.character, as.factor)

employed_df %>% 
    count(race_gender, sort = TRUE)
    
employed_df <- employed_df %>% 
    mutate(dimension = as.factor(case_when(race_gender == "TOTAL" ~ "Total",
                                 race_gender %in% c("Men", "Women") ~ "Gender",
                                 TRUE ~ "Race")))

# total employment by industry
p <- employed_df %>% 
    filter(dimension == "Total") %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) %>% 
    ggplot(aes(year, employ_n, fill = industry)) +
    geom_col(position = "dodge", alpha = 0.8) +
    labs(x = "Year",
         y = "Total # of employment",
         title = "Total # of employment in different industries",
         subtitle = "2015-2020",
         fill = "Industry") +
    scale_y_continuous(labels = comma)

plotly::ggplotly(p)    
```


```{r}
# total employment by industry and race
 employed_df %>% 
    filter(dimension == "Race") %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) %>% 
    ggplot(aes(year, employ_n, fill = race_gender)) +
    geom_col() +
    facet_wrap(~industry, scales = "free_y") +
    labs(x = "Year",
         y = "Total # of employment",
         title = "Total # of employment in different industries from 2015 to 2020",
         subtitle = "Breakdown by race",
         fill = "Race") +
    scale_y_continuous(labels = comma)

#line plot
employed_df %>% 
    filter(dimension == "Race") %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) %>%
    group_by(industry, year, race_gender) %>% 
    summarize(employ_n = sum(employ_n)) %>% 
    ggplot(aes(year, employ_n, color = race_gender)) +
    geom_line(size = 1) +
    facet_wrap(~industry, scales = "free_y") +
    labs(x = "Year",
         y = "Total # of employment",
         title = "Total # of employment in different industries from 2015 to 2020",
         subtitle = "Breakdown by Race",
         color = "Race") +
    scale_y_continuous(labels = comma)
    
```


```{r}
# total employment by industry and gender
 employed_df %>% 
    filter(dimension == "Gender") %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) %>% 
    ggplot(aes(year, employ_n, fill = race_gender)) +
    geom_col() +
    facet_wrap(~industry, scales = "free_y") +
    labs(x = "Year",
         y = "Total # of employment",
         title = "Total # of employment in different industries from 2015 to 2020",
         subtitle = "Breakdown by Gender",
         fill = "Gender") +
    scale_y_continuous(labels = comma)

# line plot
employed_df %>% 
    filter(dimension == "Gender") %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) %>%
    group_by(industry, year, race_gender) %>% 
    summarize(employ_n = sum(employ_n)) %>% 
    ggplot(aes(year, employ_n, color = race_gender)) +
    geom_line(size = 1.5) +
    facet_wrap(~industry, scales = "free_y") +
    labs(x = "Year",
         y = "Total # of employment",
         title = "Total # of employment in different industries from 2015 to 2020",
         subtitle = "Breakdown by Gender",
         color = "Gender") +
    scale_y_continuous(labels = comma)
    
```


# 2019-2020 Insights
```{r}
employed_cleaned <- employed_df %>% 
    filter(!is.na(employ_n)) %>% 
    mutate(industry = fct_reorder(industry, employ_n, sum)) 

industry_2019_2020 <- employed_cleaned %>% 
    filter(year %in% c(2019, 2020)) %>% 
    group_by(year, industry, dimension, race_gender) %>% 
    summarize(employ_n = sum(employ_n))
    
compare_2019_2020 <- industry_2019_2020 %>% 
    arrange(year) %>% 
    group_by(industry, dimension, race_gender) %>% 
    summarize(ratio = last(employ_n) / first(employ_n),
              change = ratio - 1,
              employed_2019 = first(employ_n)) %>% 
    ungroup()

compare_2019_2020 %>% 
    filter(dimension == "Total") %>% 
    mutate(industry = fct_reorder(industry, change)) %>% 
    ggplot(aes(change, industry)) + 
    geom_col() +
    scale_x_continuous(labels = percent) +
    labs(x = "Percentage change of employment",
         y = "Industry",
         title = "Percentage change of employment among different industries 2019-2020")

compare_2019_2020 %>% 
    filter(dimension == "Gender") %>% 
    mutate(industry = fct_reorder(industry, change)) %>% 
    ggplot(aes(change, industry, fill = race_gender)) + 
    geom_col(position = "dodge") +
    scale_x_continuous(labels = percent) +
    labs(x = "Percentage change of employment",
         y = "Industry",
         title = "Percentage change of employment among different industries 2019-2020",
         subtitle = "Breakdown by gender",
         fill = "Gender")

compare_2019_2020 %>% 
    filter(dimension == "Race") %>% 
    mutate(industry = fct_reorder(industry, change)) %>% 
    ggplot(aes(change, industry, fill = race_gender)) + 
    geom_col(position = "dodge") +
    scale_x_continuous(labels = percent) +
    labs(x = "Percentage change of employment",
         y = "Industry",
         title = "Percentage change of employment among different industries 2019-2020",
         subtitle = "Breakdown by race",
         fill = "Race")
```

```{r}
#lollipop graph by gender
compare_2019_2020 %>%
  filter(dimension == "Gender") %>%
  mutate(industry = fct_reorder(industry, change)) %>%
  ggplot(aes(change, industry)) +
  geom_errorbarh(aes(xmin = 0, xmax = change, color = race_gender),
                 height = 0,
                 position = position_dodge(width = .7)) +
  geom_point(aes(size = employed_2019, color = race_gender),
             position = position_dodge(width = .7)) +
  geom_vline(lty = 2, xintercept = 0) +
  scale_x_continuous(labels = percent) +
  scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  scale_size_continuous(labels = comma, guide = FALSE) +
  labs(x = "Percentage change of employment",
       y = "Industry",
       title = "Percentage change of employment among different industries 2019-2020",
       subtitle = "Breakdown by gender (Size of point represents total # of employment)",
       color = "Gender")

#lollipop graph by race
compare_2019_2020 %>%
  filter(dimension == "Race") %>%
  mutate(industry = fct_reorder(industry, change)) %>%
  ggplot(aes(change, industry)) +
  geom_errorbarh(aes(xmin = 0, xmax = change, color = race_gender),
                 height = 0,
                 position = position_dodge(width = .7)) +
  geom_point(aes(size = employed_2019, color = race_gender),
             position = position_dodge(width = .7)) +
  geom_vline(lty = 2, xintercept = 0) +
  scale_x_continuous(labels = percent) +
  scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  scale_size_continuous(labels = comma, guide = FALSE) +
  labs(x = "Percentage change of employment",
       y = "Industry",
       title = "Percentage change of employment among different industries 2019-2020",
       subtitle = "Breakdown by race (Size of point represents total # of employment)",
       color = "Race")
```



