---
title: "TidyTemplate"
date: 2021-05-04
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-05-04")

```


```{r}
library(lubridate)
water <- tt$water %>% 
  mutate(report_date = mdy(report_date)) %>% 
  filter(!country_name %in% c("Dominican Republic",
                              "Peru",
                              "Timor-Leste",
                              NA))

```

## Most common water source 
```{r}
water %>% 
    filter(!is.na(water_source)) %>% 
    count(water_source, sort = TRUE) %>% 
    mutate(water_source = fct_reorder(water_source, n)) %>% 
    ggplot(aes(n, water_source)) +
    geom_col() +
    scale_x_continuous(labels = comma) +
    labs(x = "# for each water source",
         y = "Water sources",
         title = "Most common water source")
```

## Most common water source by country
```{r}
library(tidytext)
water_source_country <- water %>%
    filter(!is.na(country_name), !is.na(water_source)) %>% 
    count(country_name, water_source, sort = TRUE) %>% 
    group_by(country_name) %>% 
    summarize(max_water_source = max(n)) %>% 
    arrange(desc(max_water_source))    # dumb way to select the top 3 observations from a group
  
count_water_source <- water %>%
    filter(!is.na(country_name), !is.na(water_source)) %>% 
    count(country_name, water_source, sort = TRUE)

water_source_country %>% 
    left_join(count_water_source, by = c("country_name", "max_water_source" = "n")) %>% 
    mutate(country_name = reorder_within(country_name, max_water_source, water_source)) %>% 
    ggplot(aes(max_water_source, country_name)) +
    geom_col() +
    scale_y_reordered() +
    facet_wrap(~water_source, scales = "free") +
    labs(x = "# for each water source",
         y = "Country",
         title = "Most common water source by country")
```


## Water system installation over time
```{r}
water %>% 
    filter(!is.na(install_year), !is.na(country_name),
           install_year <= 2021) %>% 
    count(country_name, install_year, sort = TRUE) %>% 
    ggplot(aes(install_year, n, color = country_name)) +
    geom_line() +
    geom_vline(xintercept = 2000, lty = 2, color = "black") +
    facet_wrap(~country_name, scales = "free_y", ncol = 4) +
    theme(legend.position = "none") +
    labs(x = "Year",
         y = "# of water system installation",
         title = "Number of water system installation over time",
         subtitle = "1900 - 2021 (vertical line shows year 2000)")
```


### Top 3 most used water transportation technology in each country

```{r}
water %>% 
  filter(!is.na(country_name), !is.na(water_tech)) %>% 
  count(country_name, water_tech, status_id, sort = TRUE) %>% 
  group_by(country_name) %>% 
  slice(1:3) %>%   # right way to select the top n observations from a group by country
  ungroup() %>% 
  mutate(water_tech = reorder_within(water_tech, n, country_name)) %>% 
  ggplot(aes(n, water_tech, fill = status_id)) +
  geom_col() +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  scale_fill_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown")) +
  facet_wrap(~country_name, scales = "free",
             ncol = 4) +
  labs(x = "# of water transportation technology used",
       y = "Types of water transportation technology",
       title = "Top 3 most used water transportation technology in each country",
       subtitle = "Breakdown by water availability") +
  theme(legend.position = "bottom")
```


## map

```{r}
water %>% 
  group_by(country_name) %>% 
  summarize(lon = mean(lon_deg),
            lat = mean(lat_deg)) %>% 
  ggplot(aes(lon, lat, color = country_name)) +
  geom_text(aes(label = country_name), vjust = 1, hjust= 1) +
  geom_point()

# remove dominican republic, Timor-Leste, Peru, and NA
```


```{r}
# show the distribution on world map
water %>% 
  sample_n(50000) %>% 
  ggplot(aes(lon_deg, lat_deg, color = country_name)) +
  geom_point() +
  borders("world") +
  theme(legend.position = "bottom")
```


```{r}
# show African map
library(countrycode)
africa_map <- map_data("world") %>% 
  as_tibble() %>% 
  mutate(continent = countrycode(region, "country.name", "continent")) %>% 
  filter(continent == "Africa")

water %>% 
  sample_n(50000) %>% 
  ggplot(aes(lon_deg, lat_deg)) +
  geom_polygon(aes(long, lat, group = group), data = africa_map,
               fill = "white",
               color = "grey") +
  geom_point(size = 0.1, alpha = 0.1) +
  xlim(c(-25, 50)) +
  ylim(c(-30, 40))

```


```{r}
# use google API to show the distribution of water source among african countries
library(ggmap)
register_google(key = "AIzaSyB2c-7a4NoG5qadKfP2fbAMDTMthd05RYc")

africa_water <- water %>% 
  filter(!is.na(water_source)) %>% 
  mutate(water_source = fct_lump(water_source, 5))
  
ggmap(get_map("Africa", zoom = 3)) +
  geom_point(aes(lon_deg, lat_deg, color = water_source),
             data = africa_water %>% sample_n(50000),
             size = 0.1,
             alpha = 0.1) +
  facet_wrap(~water_source) +
  theme(legend.position = "none")
  
```


```{r}
# Show only Uganda
library(ggthemes)
uganda <- water %>% 
  filter(!is.na(water_source),
         country_name == "Uganda") %>% 
  mutate(water_source = fct_lump(water_source, 5),
         water_source = fct_reorder(water_source, water_source,
                                    length, .desc = TRUE))
# use borders
uganda %>% 
  ggplot(aes(lon_deg, lat_deg, color = status_id))+
  geom_point(size = 0.1, alpha = 0.2) +
  borders("world", regions = "Uganda") +
  xlim(c(29, 35)) +
  ylim(c(-2, 5)) +
  scale_color_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown"),
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  theme_map() +
  labs(title = "Water availability in Uganda")
```


```{r}
# facet on water source
uganda %>% 
  ggplot(aes(lon_deg, lat_deg, color = water_source))+
  geom_point(size = 0.1, alpha = 0.2, show.legend = FALSE) +
  borders("world", regions = "Uganda") +
  xlim(c(29, 35)) +
  ylim(c(-2, 5)) +
  facet_wrap(~water_source) +
  theme_map() +
  labs(title = "Water sources in Uganda")
```

```{r}
# use stamenmap

ggmap(get_stamenmap(bbox = c(left = 29, bottom = -2,
                             right = 35, top = 5), zoom = 8)) +
  geom_point(aes(lon_deg, lat_deg, color = water_source),
             size = 0.1, alpha = 0.2, show.legend = FALSE,
             data = uganda) +
  facet_wrap(~water_source) +
  theme_map() +
  labs(title = "Water sources in Uganda")
  
```


```{r}
# animation Uganda data
library(gganimate)

uganda_animate <- uganda %>%
  mutate(install_year = ifelse(install_year > 2021, NA, install_year)) %>% 
  filter(!is.na(install_year)) %>% 
  sample_n(10000) %>% 
  mutate(year = map(install_year, ~ seq(., 2021))) %>% 
  unnest(year)

```

```{r}
# Uganda water availability over time (borders)
uganda_animate %>% 
  ggplot(aes(lon_deg, lat_deg, color = status_id))+
  geom_point(size = 0.5, alpha = 0.5) +
  borders("world", regions = "Uganda") +
  xlim(c(29, 35)) +
  ylim(c(-2, 5)) +
  scale_color_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown"),
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  labs(title = "Water availability in Uganda over time: { current_frame }")
```


```{r}
# Uganda water source change over time (stamenmap)
ggmap(get_stamenmap(bbox = c(left = 29, bottom = -2,
                             right = 35, top = 5), zoom = 8)) +
  geom_point(aes(lon_deg, lat_deg, color = water_source),
             size = 0.5, alpha = 0.5,
             data = uganda_animate) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  scale_color_discrete(name = "Water sources",
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  labs(title = "Change of water sources in Uganda over time: { current_frame }")
```

```{r}
# google api same map
ggmap(get_map("Uganda", zoom = 7)) +
  geom_point(aes(lon_deg, lat_deg, color = water_source),
             size = 0.5, alpha = 0.5,
             data = uganda_animate) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  scale_color_discrete(name = "Water sources",
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  labs(title = "Change of water sources in Uganda over time: { current_frame }")
```


```{r}
# Africa animate data
africa_animate <- water %>% 
  filter(!is.na(water_source)) %>% 
  mutate(water_source = fct_lump(water_source, 5),
         water_source = fct_reorder(water_source, water_source,
                                    length, .desc = TRUE)) %>% 
  mutate(install_year = ifelse(install_year > 2021, NA, install_year)) %>% 
  filter(!is.na(install_year)) %>% 
  sample_n(10000) %>% 
  mutate(year = map(install_year, ~ seq(., 2021))) %>% 
  unnest(year)


```


```{r}
# water availability in Africa (borders)
africa_animate %>% 
  ggplot(aes(lon_deg, lat_deg, color = status_id))+
  geom_polygon(aes(long, lat, group = group), data = africa_map,
               fill = "white",
               color = "grey") +
  geom_point(size = 0.5, alpha = 0.5) +
  xlim(c(-25, 50)) +
  ylim(c(-30, 40)) +
  scale_color_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown"),
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  labs(title = "Water availability in Africa over time: { current_frame }")

```


```{r}
# water availability in Africa (stamenmap)
ggmap(get_stamenmap(bbox = c(left = -22, bottom = -36,
                             right = 53, top = 40), zoom = 4)) +
  geom_point(aes(lon_deg, lat_deg, color = status_id),
             size = 0.8, alpha = 0.7,
             data = africa_animate) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  scale_color_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown"),
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  labs(title = "Water availability in Africa over time: { current_frame }")
```


```{r}
# water availability in Africa (google map)
ggmap(get_map("Africa", zoom = 3)) +
  geom_point(aes(lon_deg, lat_deg, color = status_id),
             size = 0.5, alpha = 0.5,
             data = africa_animate) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  scale_color_discrete(name = "Water availability",
                      limits = c("y", "n", "u"),
                      labels = c("Yes", "No", "Unknown"),
                      guide = guide_legend(override.aes = 
                                              list(size = 2, alpha = 1))) +
  transition_manual(year) +
  theme_map() +
  theme(legend.position = "right") +
  labs(title = "Water availability in Africa over time: { current_frame }")
```

