---
title: "TidyTemplate"
date: 2021-05-17
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(ggthemes)
library(tidymodels)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}
library(lubridate)
tt <- tt_load("2021-05-18")
survey <- tt$survey
tidy_survey <- survey %>% 
    mutate(timestamp = as_datetime(timestamp, format = "%m/%d/%Y %H:%M:%S")) %>% 
    mutate(across(.cols = annual_salary:other_monetary_comp, as.numeric),
           across(.cols = c(how_old_are_you:industry, currency,
                            state, overall_years_of_professional_experience:
                                race), as.factor),
           country = str_replace(country, "^[Uu]\\.*[Ss]\\.*([Aa])*\\.*",
                                 "United States"),
           country = str_replace(country, "(The )*([Uu].+) [Ss].+",
                                 "United States"),
           country = str_replace(country, "America", "United States"),
           country = str_replace(country, "america", "United States"),
           country = str_replace(country, "The US", "United States"),
           country = str_replace(country, "California", "United States"),
           country = str_replace(country, "San Francisco", "United States"),
           country = str_replace(country, "U. S", "United States"),
           country = str_replace(country, "^[Uu][Kk]",
                                 "United Kingdom"),
           country = str_replace(country, "^Can.+", "Canada"),
           country = str_replace(country, "^can.+", "Canada"),
           country = str_replace(country, "CANADA", "Canada"),)
```

# How long the survey last?

```{r}
last(tidy_survey$timestamp) - first(tidy_survey$timestamp)
```

# US respondents' industry

```{r}
us_only <- tidy_survey %>% 
    filter(country == "United States",
           currency != "Other") %>%
        mutate(how_old_are_you = factor(how_old_are_you,
                                    levels = c("under 18",
                                               "18-24", "25-34",
                                               "35-44", "45-54",
                                               "55-64", "65 or over"))) %>% 
    mutate(annual_salary = case_when(
                                     currency == "EUR" ~ annual_salary * 1.22,
                                     currency == "JPY" ~ annual_salary * 0.009,
                                     currency == "GBP" ~ annual_salary * 1.42,
                                     currency == "CHF" ~ annual_salary * 1.11,
                                     currency == "CAD" ~ annual_salary * 0.83,
                                     currency == "AUD/NZD" ~ annual_salary * 0.93,
                                     currency == "ZAR" ~ annual_salary * 0.07,
                                     currency == "HKD" ~ annual_salary * 0.13,
                                     currency == "SEK" ~ annual_salary * 0.12,
                                     TRUE ~ annual_salary)) %>% 
    mutate(other_monetary_comp = ifelse(is.na(other_monetary_comp),
                                        0, other_monetary_comp)) %>% 
    mutate(total_income = annual_salary + other_monetary_comp)

us_only %>% 
    count(industry, sort = TRUE) %>% 
    mutate(industry = fct_reorder(industry, n)) %>% 
    slice(1:10) %>% 
    ggplot(aes(n, industry)) +
    geom_col(aes(fill = industry, color = industry),
             show.legend = FALSE) + 
    geom_text(aes(label = n), hjust = -0.2) +
    labs(x = "# of respondents",
         y = "",
         title = "Top 10 industry U.S. respondents working in")
```

# US individuals with highest total income

```{r}
us_only %>% 
    filter(currency != "Other") %>% 
    slice_max(total_income, n = 20, with_ties = FALSE)
```


# Individuals with the top 3 highest total income in different industries

```{r}
library(tidytext)
library(ggrepel)
us_only %>% 
    filter(industry %in% c("Accounting, Banking & Finance",
                           "Agriculture or Forestry",
                           "Art & Design",
                           "Business or Consulting",
                           "Computing or Tech",
                           "Education (Primary/Secondary)",
                           "Education (Higher Education)",
                           "Engineering or Manufacturing",
                           "Entertainment",
                           "Government and Public Administration",
                           "Health care",
                           "Hospitality & Events",
                           "Insurance",
                           "Law",
                           "Law Enforcement & Security",
                           "Leisure, Sport & Tourism",
                           "Marketing, Advertising & PR",
                           "Media & Digital",
                           "Nonprofits",
                           "Property or Construction",
                           "Recruitment or HR",
                           "Retail",
                           "Sales",
                           "Social Work",
                           "Transport or Logistics",
                           "Utilities & Telecommunications")) %>% 
    group_by(industry) %>% 
    slice_max(total_income, n = 3, with_ties = FALSE) %>%
    mutate(job_title = reorder_within(job_title, total_income, industry)) %>% 
    ggplot(aes(total_income, job_title)) +
    scale_y_reordered() +
    scale_x_continuous(labels = dollar) +
    geom_col(aes(fill = gender), position = "dodge") +
    geom_text(aes(label = highest_level_of_education_completed), hjust = 1,
              size = 2.5) +
    facet_wrap(~industry, scales = "free", ncol = 3) +
    labs(x = "Total income",
         y = "",
         title = "Individuals with the top 3 highest total income in different industries in the U.S.",
         subtitle = "Breakdown by job title and gender",
         fill = "Gender") +
     ggthemes::theme_clean()
```


# Total income by age group in different industries

```{r}
us_only %>% 
    filter(industry %in% c("Accounting, Banking & Finance",
                           "Agriculture or Forestry",
                           "Art & Design",
                           "Business or Consulting",
                           "Computing or Tech",
                           "Education (Primary/Secondary)",
                           "Education (Higher Education)",
                           "Engineering or Manufacturing",
                           "Entertainment",
                           "Government and Public Administration",
                           "Health care",
                           "Hospitality & Events",
                           "Insurance",
                           "Law",
                           "Law Enforcement & Security",
                           "Leisure, Sport & Tourism",
                           "Marketing, Advertising & PR",
                           "Media & Digital",
                           "Nonprofits",
                           "Property or Construction",
                           "Recruitment or HR",
                           "Retail",
                           "Sales",
                           "Social Work",
                           "Transport or Logistics",
                           "Utilities & Telecommunications")) %>% 
    filter(!is.na(highest_level_of_education_completed)) %>% 
    ggplot(aes(total_income, how_old_are_you)) +
    ggridges::geom_density_ridges(aes(fill = how_old_are_you),
                                  show.legend = FALSE) +
    # geom_boxplot(aes(fill = how_old_are_you, color = how_old_are_you),
    #              alpha = 0.7, show.legend = FALSE) +
    scale_x_continuous(labels = dollar) +
    facet_wrap(~industry, scales = "free_x", ncol = 4) +
    labs(x = "Total income",
         y = "",
         color = "Education",
         title = "Total income varied by age group in different industries") +
    ggthemes::theme_clean()
```


# PhD total income by age group in different industries 

```{r}
us_only %>% 
    filter(industry %in% c("Accounting, Banking & Finance",
                           "Agriculture or Forestry",
                           "Art & Design",
                           "Business or Consulting",
                           "Computing or Tech",
                           "Education (Primary/Secondary)",
                           "Education (Higher Education)",
                           "Engineering or Manufacturing",
                           "Entertainment",
                           "Government and Public Administration",
                           "Health care",
                           "Hospitality & Events",
                           "Insurance",
                           "Law",
                           "Law Enforcement & Security",
                           "Leisure, Sport & Tourism",
                           "Marketing, Advertising & PR",
                           "Media & Digital",
                           "Nonprofits",
                           "Property or Construction",
                           "Recruitment or HR",
                           "Retail",
                           "Sales",
                           "Social Work",
                           "Transport or Logistics",
                           "Utilities & Telecommunications")) %>% 
    filter(highest_level_of_education_completed == "PhD") %>% 
    ggplot(aes(total_income, how_old_are_you)) +
    geom_point(alpha = 0.4) +
    geom_boxplot(aes(fill = how_old_are_you,
                     color = how_old_are_you),
                 alpha = 0.8,
                 show.legend = FALSE) +
    scale_x_continuous(labels = dollar) +
    facet_wrap(~industry, scales = "free_x", ncol = 4) +
    labs(x = "Total income",
         y = "",
         title = "PhDs' total income varied by age group in different industries") +
    ggthemes::theme_clean()
```

# Highest paid individual PhD by positions in government

```{r}
us_only %>% 
    filter(highest_level_of_education_completed == "PhD",
           industry == "Government and Public Administration") %>% 
    group_by(job_title) %>% 
    slice_max(annual_salary, n = 1, with_ties = FALSE) %>% 
    ungroup() %>% 
    mutate(job_title = fct_reorder(job_title, annual_salary)) %>% 
    ggplot(aes(annual_salary, job_title)) +
    geom_col(aes(fill = gender), position = "dodge") +
    scale_x_continuous(labels = dollar) +
    labs(x = "Annual Salary",
         y = "",
         title = "Highest paid individual PhD by positions in government",
         fill = "Gender") +
    ggthemes::theme_clean(base_family = "serif")

# Remove that one senior policy adviser
us_only %>% 
    filter(highest_level_of_education_completed == "PhD",
           industry == "Government and Public Administration",
           job_title != "Senior Policy Advisor") %>% 
    group_by(job_title) %>% 
    slice_max(annual_salary, n = 1, with_ties = FALSE) %>% 
    ungroup() %>% 
    mutate(job_title = fct_reorder(job_title, annual_salary)) %>% 
    ggplot(aes(annual_salary, job_title)) +
    geom_col(aes(fill = gender), position = "dodge") +
    scale_x_continuous(labels = dollar) +
    labs(x = "Annual Salary",
         y = "",
         title = "Highest paid individual PhD by positions in government",
         fill = "Gender") +
    ggthemes::theme_clean(base_family = "serif")
```

# Median PhD annual salary by positions in governemnt

```{r}
us_only %>% 
    filter(highest_level_of_education_completed == "PhD",
           industry == "Government and Public Administration") %>% 
    group_by(job_title) %>% 
    summarize(median_salary = median(annual_salary)) %>% 
    ungroup() %>% 
    mutate(job_title = fct_reorder(job_title, median_salary)) %>% 
    ggplot(aes(median_salary, job_title)) +
    geom_col() +
    scale_x_continuous(labels = dollar) +
    labs(x = "Median Salary",
         y = "",
         title = "Median PhD annual salary by positions in governemnt") +
    ggthemes::theme_clean(base_family = "serif")

# Remove that one senior policy adviser
us_only %>% 
    filter(highest_level_of_education_completed == "PhD",
           industry == "Government and Public Administration",
           job_title != "Senior Policy Advisor") %>% 
    group_by(job_title) %>% 
    summarize(median_salary = median(annual_salary)) %>% 
    ungroup() %>% 
    mutate(job_title = fct_reorder(job_title, median_salary)) %>% 
    ggplot(aes(median_salary, job_title)) +
    geom_col() +
    scale_x_continuous(labels = dollar) +
    labs(x = "Median Salary",
         y = "",
         title = "Median PhD annual salary by positions in governemnt") +
    ggthemes::theme_clean(base_family = "serif")
```

# Median salary for data scientist or data analyst in diffrent industries

```{r}
us_only %>% 
    filter(job_title %in% c("Data Scientist", "Data Analyst")) %>% 
    group_by(industry, job_title) %>% 
    summarize(median_salary = median(annual_salary)) %>% 
    ungroup() %>% 
    mutate(industry = reorder_within(industry, median_salary, job_title)) %>% 
    ggplot(aes(median_salary, industry, fill = job_title,
               color = job_title, alpha = 0.7)) +
    geom_col(show.legend = FALSE) +
    scale_x_continuous(labels = dollar) +
    scale_y_reordered() +
    facet_wrap(~job_title, scales = "free_y") +
    labs(x = "Median Salary",
         y = "",
         title = "Median salary for data scientist or data analyst in diffrent industries") +
    ggthemes::theme_clean(base_family = "serif")
```

# Highest paid individual data scientist or data analyst in diffrent industries

```{r}
us_only %>% 
    filter(job_title %in% c("Data Scientist", "Data Analyst")) %>% 
    group_by(industry) %>% 
    slice_max(annual_salary, n = 1, with_ties = FALSE) %>% 
    ungroup() %>% 
    mutate(industry = reorder_within(industry, annual_salary, job_title)) %>% 
    ggplot(aes(annual_salary, industry)) +
    geom_col(aes(fill = job_title,
                 color = job_title, alpha = 0.7),
             show.legend = FALSE) +
    geom_text(aes(label = paste0(gender, " (",
                                 highest_level_of_education_completed,
                                 ")")), hjust = 1,
              size = 3.5, family = "serif") +
    scale_x_continuous(labels = dollar) +
    scale_y_reordered() +
    facet_wrap(~job_title, scales = "free_y") +
    labs(x = "Annual Salary",
         y = "",
         title = "Highest paid individual data scientist or data analyst in diffrent industries") +
    ggthemes::theme_clean(base_family = "serif")
```


# Does the type of professional experience matter?

```{r}
us_only %>% 
    pivot_longer(cols = overall_years_of_professional_experience:years_of_experience_in_field,
                 names_to = "experience_type",
                 values_to = "year_of_experience") %>% 
    mutate(year_of_experience = fct_reorder(year_of_experience,
                                            annual_salary)) %>% 
    ggplot(aes(annual_salary, year_of_experience)) +
    geom_boxplot(aes(
                     color = year_of_experience),
                     show.legend = FALSE) +
    facet_wrap(~experience_type) +
    scale_x_log10(label = dollar) +
    labs(x = "Annual salary (log transformed)",
         y = "",
         title = "Annual salary by year of experience",
         subtitle = "Overall experience vs. experience in field") +
    ggthemes::theme_clean(base_family = "serif")
```

As the year of experience increases, the median annual salary increases. However, this trend disappears when the year of experience reaches 41 years or more.

# Does the gender matter?

```{r}
us_only %>% 
    filter(!is.na(gender),
           !gender %in% c("Prefer not to answer",
                         "Other or prefer not to answer")) %>% 
    mutate(gender = fct_reorder(gender, annual_salary)) %>% 
    ggplot(aes(annual_salary, gender)) +
    geom_boxplot(aes(color = gender),
                     show.legend = FALSE) +
    facet_wrap(~how_old_are_you, scales = "free",
               ncol = 2) +
    scale_x_log10(label = dollar) +
    labs(x = "Annual salary (log transformed)",
         y = "",
         title = "Annual salary discrepencies by gender and age") +
    ggthemes::theme_clean(base_family = "serif")
```

Gender definitely matters and man tends to make more across all age groups except for people under 18.
