---
title: "TidyTemplate"
date: 2021-07-14
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly \#TidyTuesday event! Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data. While the dataset will be "tamed", it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format. The goal of TidyTuesday is to apply your R skills, get feedback, explore other's work, and connect with the greater \#RStats community! As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(DataExplorer)
library(GGally)
theme_set(theme_light())
doParallel::registerDoParallel(cores = 8)
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}
# Solve the NULL as NA issue in the original data by adding na = argument
scoobydoo <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-13/scoobydoo.csv', na = c("", "NA", "NULL")) %>% 
    mutate(number_of_snacks = parse_number(number_of_snacks)) %>%
    filter(
           index != 293,
           index != 313,
           index != 393,
           index != 394,
           index != 404,
           index != 408,
           index != 410,
           index != 411,
           index != 417,
           index != 576)
    # # problem with monster_name column (6 categories) while other columns containing "monster_" 
    # has only 5 categories, which makes the use of seperate_rows problematic

scoobydoo[60, 11] <- "Mamba Wamba,Zombie Lila,Mamba Zombie" # fix extra "," in index 60 monster_name
scoobydoo[84, 12] <- "Female,Male" # fix extra "," in index 60 monster_gender
scoobydoo[28, 47] <- "House Keeping Couple,House Keeping Couple" # fix index 28 culprit_name only one "House Keeping Couple"
scoobydoo <- scoobydoo %>% 
    mutate(culprit_name = if_else(index == 602, "Waitress,Cook,Janitor",
                                  culprit_name))# fix index 602 culprit_name "." instead of "," before Janitor
```

```{r}
mn_sep <- scoobydoo %>%
    separate_rows(monster_name, sep = ",")

mg_sep <- scoobydoo %>%
    separate_rows(monster_gender, sep = ",")

mt_sep <- scoobydoo %>%
    separate_rows(monster_type, sep = ",")

mst_sep <- scoobydoo %>%
    separate_rows(monster_subtype, sep = ",")

ms_sep <- scoobydoo %>%
    separate_rows(monster_species, sep = ",")

cn_sep <- scoobydoo %>%
    separate_rows(culprit_name, sep = ",")

cg_sep <- scoobydoo %>% 
    separate_rows(culprit_gender, sep = ",")
```


```{r}
mn_sep %>% select(index, monster_name, monster_gender) %>% mutate(id = row_number()) 
mg_sep %>% select(index, monster_name, monster_gender) %>% mutate(id = row_number()) 

mn_sep %>% select(index, monster_name, monster_gender) %>% mutate(id = row_number()) %>% 
    anti_join(mg_sep %>% select(index, monster_name, monster_gender) %>% 
                  mutate(id = row_number()),
              by = "id")

# index 293 monster_gender should have 6 Males instead of 5 given that there are six monster names for index 293

# index 84 monster_gender has an extra ","
# index 393 monster_name has more categories than others
#index 394 monster_name has less categories than others
# index 404 monster_name has more categories than others
# index 417 monster_name has more categories than others
```


```{r}
mn_sep %>% select(index, monster_name, monster_type) %>% mutate(id = row_number()) 
mt_sep %>% select(index, monster_name, monster_type) %>% mutate(id = row_number()) 

mn_sep %>% select(index, monster_name, monster_type) %>% mutate(id = row_number()) %>% 
    anti_join(mt_sep %>% select(index, monster_name, monster_type) %>% 
                  mutate(id = row_number()),
              by = "id")
# index 60 monster_name has an extra "," in the end.
# index 313 monster_name has only 17 categories while others have more categories.
# index 408 monster_name has less categories 
# index 576 monster_name has less categories
```

```{r}
mn_sep %>% select(index, monster_name, monster_species) %>% mutate(id = row_number()) 
ms_sep %>% select(index, monster_name, monster_species) %>% mutate(id = row_number())

mn_sep %>% select(index, monster_name, monster_species) %>% mutate(id = row_number()) %>% 
    anti_join(mt_sep %>% select(index, monster_name, monster_species) %>% 
                  mutate(id = row_number()),
              by = "id")

# index 410 monster_name has less categories 
# index 411 monster_name has more categories 
```



## Cleaned data
```{r}
scoobydoo <- scoobydoo %>% 
    separate_rows(monster_name:monster_species,
                  sep = ",")
```

## EDA

```{r}
scoobydoo %>% 
    skimr::skim()
```

## Correlation of all numeric variables

```{r}
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    na.omit() %>% 
    plot_correlation()
```

* imdb is primarily associated with `engagement`, `run_time`, `monster_amount`, `suspects_amount`, `jinkies`, `my_glasses`and `rooby_rooby_roo`.

## Visualize scatterplots between imdb and all numeric variables
```{r}
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    relocate(imdb, .after = rooby_rooby_roo) %>% 
    ggpairs(c(1:8, 18)) # look at the first 8 variables
```

* `engagement`, `run_time`, `monster_amount` are potential predictors of imdb
(`suspects_amount` is highly correlated with `run_time` and is not important in
this case)

```{r}
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    relocate(imdb, .after = rooby_rooby_roo) %>% 
    ggpairs(10:18) # look at the last 8 variables
```
* `jinkies`, `my_glasses`, and `rooby_rooby_roo` are potential predictors of imdb

### Examine important numeric variables carefully

```{r}
# Write a function to construct scatterplot between imdb and other numeric
#variables
plot_scatter <- function (tbl, variable){
    tbl %>% 
        ggplot(aes({{ variable }}, imdb)) +
        geom_point() +
        geom_smooth(method = "lm", color = "red") +
        geom_smooth(method = "loess") 
}

# engagement
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(engagement) +
    scale_x_log10() # engagement seems to have a quadratic effect on imdb score

# run_time
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(run_time) # the longer the run time, the lowerer the imdb score

# monster_amount
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(monster_amount) # imdb score seems to decrease when monster amount increases

# jinkies
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(jinkies) # imdb score seems to decrease when jinkies increases

# my_glasses
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(my_glasses) # imdb score seems to decrease when my_glasses increases

# rooby_rooby_doo
scoobydoo %>% 
    select_if(is.numeric) %>% 
    select(-index) %>% 
    plot_scatter(rooby_rooby_roo) # imdb score seems to decrease when rooby increases
```

## Relationships between imdb and all logical variables

```{r}
# Density plots
scoobydoo %>% 
    select_if(is.logical) %>% 
    bind_cols(scoobydoo[6]) %>% 
    relocate(imdb) %>% 
    pivot_longer(monster_real:blue_falcon,
                 names_to = "variable",
                 values_to = "value") %>% 
    na.omit() %>% 
    ggplot(aes(imdb, color = value)) +
    geom_density() +
    facet_wrap(~variable, scales = "free", ncol = 5) +
    theme(legend.position = "bottom") +
    labs(x = "imdb scores",
         y = "Density",
         color = "Status",
         title = "Distribution of imdb scores",
         subtitle = "Breakdown by status change (TRUE or FALSE) in different characteristics")
```

* Based on the density plot, most status change don't have impact on imdb scores.
However, `blue_falcon`, `caught_not`, and `snack_shaggy` seem to matter.

```{r}
# boxplots
scoobydoo %>% 
    select_if(is.logical) %>% 
    bind_cols(scoobydoo[6]) %>% 
    relocate(imdb) %>% 
    pivot_longer(monster_real:blue_falcon,
                 names_to = "variable",
                 values_to = "value") %>% 
    na.omit() %>% 
    ggplot(aes(imdb, value, fill = value)) +
    geom_boxplot(alpha = 0.6) +
    facet_wrap(~variable, scales = "free", ncol = 5) +
    theme(legend.position = "bottom") +
    labs(x = "imdb scores",
         fill = "Status",
         title = "Distribution of imdb scores",
         subtitle = "Breakdown by status change (TRUE or FALSE) in different characteristics")
```

* `monster_real` seems to matter as well.

## Relationships between imdb and all character variables

```{r}
# Find the number of unique categories in each character variable
scoobydoo %>% 
    select_if(is.character) %>% 
    map_df(~ count(data.frame(x = .), x, sort = TRUE), .id = "variable_name") %>% 
    as_tibble() %>% 
    group_by(variable_name) %>% 
    summarize(unique_categories = n()) %>% 
    arrange(desc(unique_categories)) %>% 
    print(n = Inf)
```



```{r}
# Or achieve it through the following way
scoobydoo %>% 
    select_if(is.character) %>% 
    map_dbl(~ length(unique(.x)))
```

```{r}
# Count of unique categories for each variable
scoobydoo %>% 
    select_if(is.character) %>% 
    map_df(~ count(data.frame(x = .), x, sort = TRUE), .id = "variable_name") %>% 
    as_tibble() %>% 
    split(.$variable_name)
```


```{r}

plot_col <- function (tbl, variable){
    if (is.character(pull(tbl, {{ variable }}))) {
        tbl <- tbl %>% 
        mutate({{ variable }} := fct_lump({{ variable }}, 10),
               {{ variable }} := fct_reorder({{ variable }}, imdb))
    }
    tbl %>% 
        ggplot(aes(imdb, {{ variable }})) +
        geom_col()
}

# series_name
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(series_name) # seem to have effect

# network
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(network) # seem to matter

# season
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(season) # seem to matter

# title
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(title) # doesn't matter

# format
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(format) # seem to matter

# monster_name
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(monster_name) # doesn't matter

# monster_gender
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(monster_gender) # matters a lot

# monster_type
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(monster_type) # seem to matter

# monster_subtype
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(monster_subtype) # doesn't matter much

# monster_species
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(monster_species) # human vs. others matter

# setting_terrain
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(setting_terrain) # urban, rural, vs. others

# setting_country_state
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(setting_country_state) # United States, California, vs. others.

# culprit_name
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(culprit_name) # doesn't matter

# culprit_gender
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(culprit_gender) # probably doesn't matter much and alos it's messed up

# motive
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(motive) # matters

# if_it_wasnt_for
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(if_it_wasnt_for) # doesn't matter

# and_that
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(and_that) # doesn't matter

# fred_va
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(fred_va) # Frank Welker vs. others

# daphnie_va
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(daphnie_va) # Grey DeLisle or Heather North vs. others

# velma_va
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(velma_va) # Mindy Cohn vs. others

# shaggy_va
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(shaggy_va) # Casey kasem vs. others

# scooby_va
scoobydoo %>% 
    select_if(is.character) %>% 
    bind_cols(scoobydoo %>% select(imdb)) %>% 
    relocate(imdb) %>% 
    plot_col(scooby_va) # Don Messick or Frank Welker vs. others
```

`and_that`, `if_it_wasnt_for`, `culprit_gender`, `culprit_name`, `monster_subtype`, `monster_name`,
`title` will be excluded from the model due to no/lower impacts on imdb score.

### Does the time matter?

```{r}
scoobydoo %>% 
    ggplot(aes(date_aired, imdb, color = season)) +
    geom_line() +
    facet_wrap(~season) +
    labs(x = "Aired date",
         y = "imdb score",
         title = "imdb score over time",
         subtitle = "Breakdown by seasons",
         color = "Season") +
    theme(legend.position = "bottom")
```

Overall, it looks like some random noise between aired date and imdb score.

```{r}
# use decade instead of date
library(lubridate)
scoobydoo %>% 
    ggplot(aes(10*(year(date_aired) %/%10), imdb)) +
    geom_line() +
    geom_smooth(method = "loess") +
    labs(x = "Aired date by decade",
         y = "imdb score",
         title = "imdb score over decades") 
```

When switch to decade, it seems that there is a nonlinear relationship between imbd score and aired decade.



## Modelling

### Train, test, and validation sets

```{r}
set.seed(121)
library(tidymodels)


scoobydoo_model <- scoobydoo %>% 
    select(index, imdb, engagement, run_time, monster_amount, jinkies, my_glasses, rooby_rooby_roo, blue_falcon, caught_not, snack_shaggy, monster_real, series_name:season, format, monster_gender, monster_type, monster_species, setting_terrain, setting_country_state, motive, fred_va:scooby_va, date_aired) %>% 
    mutate_if(is.character, as.factor)

scoobydoo_split <- initial_split(scoobydoo_model, strata = imdb)
scoobydoo_train <- training(scoobydoo_split)
scoobydoo_test <- testing(scoobydoo_split)

set.seed(123)
scoobydoo_fold <- vfold_cv(scoobydoo_train, v = 10)
```


### Feature engineering

```{r}
scoobydoo_rec <- recipe(imdb ~ ., data = scoobydoo_train) %>% 
    update_role(index, new_role = "id") %>% 
    step_mutate(date_aired = 10 * (year(date_aired) %/% 10),
                monster_real = factor(monster_real),
                snack_shaggy = factor(snack_shaggy)) %>% 
    step_other(all_nominal_predictors(), threshold = 0.1) %>% 
    step_impute_median(all_numeric_predictors(), all_outcomes()) %>% 
    step_impute_mode(all_nominal_predictors()) %>% 
    step_log(engagement, base = 10, offset = 1) %>% 
    step_unknown(snack_shaggy, new_level = "unknown") %>%
    step_unknown(monster_real, new_level = "unknown") %>%# since snack_shaggy and monster_real still have missing values even after imputation, I set the missing value as unknown.
    step_dummy(all_nominal_predictors()) %>% 
    step_ns(engagement, deg_free = tune("engagement df")) %>% 
    step_ns(run_time, deg_free =  tune("run_time df")) %>% 
    step_ns(date_aired, deg_free = tune("date_aired df"))

```

### linear model

```{r}
# model
scoobydoo_linear_model <- linear_reg() %>% 
    set_engine("lm") %>% 
    set_mode("regression")
```

```{r}
# workflow
scoobydoo_linear_wf <- workflow() %>% 
    add_recipe(scoobydoo_rec) %>% 
    add_model(scoobydoo_linear_model)
```

### Tuning linear model

```{r}
set.seed(2021)

scoobydoo_param <- scoobydoo_linear_wf %>% 
    parameters()

scoobydoo_linear_res <- scoobydoo_linear_wf %>% 
    tune_grid(scoobydoo_fold,
              grid = grid_regular(scoobydoo_param, levels = 5))

autoplot(scoobydoo_linear_res)

scoobydoo_linear_res %>% 
    collect_metrics() %>% 
    arrange(mean)
```

* The result clearly show that adding splines to date_aired variable is not helpful, while spline effects do exist for engagement and run_time at 15. Thus, we can update the recipe to train an xgboost model.

### xgboost model

#### update recipe for xgboost model

```{r}
scoobydoo_rec_xg <- recipe(imdb ~ ., data = scoobydoo_train) %>% 
    update_role(index, new_role = "id") %>% 
    step_mutate(date_aired = 10 * (year(date_aired) %/% 10),
                monster_real = factor(monster_real),
                snack_shaggy = factor(snack_shaggy)) %>% 
    step_other(all_nominal_predictors(), threshold = 0.1) %>% 
    step_impute_median(all_numeric_predictors(), all_outcomes()) %>% 
    step_impute_mode(all_nominal_predictors()) %>% 
    step_log(engagement, base = 10, offset = 1) %>% 
    step_unknown(snack_shaggy, new_level = "unknown") %>%
    step_unknown(monster_real, new_level = "unknown") %>%# since snack_shaggy and monster_real still have missing values even after imputation, I set the missing value as unknown.
    step_dummy(all_nominal_predictors()) %>% 
    step_ns(engagement, deg_free = 15) %>% 
    step_ns(run_time, deg_free =  15)

scoobydoo_rec_xg %>% 
    prep() %>% 
    juice()
```

```{r}
scoobydoo_spec_xg <- boost_tree(mtry = tune(),
                                trees = tune(),
                                tree_depth = tune(),
                                learn_rate = tune()) %>% 
    set_engine("xgboost") %>% 
    set_mode("regression")

scoobydoo_xg_wf <- workflow() %>% 
    add_recipe(scoobydoo_rec_xg) %>% 
    add_model(scoobydoo_spec_xg)
```


### Tuning xgboost model

```{r}

# use latin_hypercude

set.seed(007)

scoobydoo_xg_res <- scoobydoo_xg_wf %>% 
    tune_grid(scoobydoo_fold,
              grid = grid_latin_hypercube(finalize(mtry(), scoobydoo_train),
                                  trees(), tree_depth(),
                                  learn_rate(), size = 100))

autoplot(scoobydoo_xg_res)

scoobydoo_xg_res %>% 
    collect_metrics() %>% 
    arrange(mean)
```

* Not the best result

```{r}
# using crossing
set.seed(2021)
scoobydoo_xg_res <- scoobydoo_xg_wf %>% 
    tune_grid(scoobydoo_fold,
              grid = crossing(mtry = seq(10, 35, 5),
                              trees = seq(300, 1000, 100),
                              tree_depth = seq(5, 15, 5),
                              learn_rate = c(0.01, 0.02)),
              control = control_grid(save_pred = TRUE,
                                     save_workflow = TRUE))

autoplot(scoobydoo_xg_res)

scoobydoo_xg_res %>% 
    collect_metrics() %>% 
    arrange(mean)
```

### Explore results

```{r}
show_best(scoobydoo_xg_res, "rmse")
show_best(scoobydoo_xg_res, "rsq")
```

* The best rmse for training data tuned by xgboost model is `show_best(scoobydoo_xg_res, "rmse")`.

### Finalize xgboost workflow
```{r}
final_xg_wf <- finalize_workflow(scoobydoo_xg_wf,
                      select_best(scoobydoo_xg_res, "rmse"))
```


### Important variables
```{r}
library(vip)
final_xg_wf %>% 
    fit(data = scoobydoo_train) %>% 
    pull_workflow_fit() %>% 
    vip(geom = "col")
```

### Fitting on test data

```{r}
final_xg_res <- last_fit(final_xg_wf, scoobydoo_split)
final_xg_res %>% 
    collect_metrics()
```

* The best rmse for testing data tuned by xgboost model is `show_best(final_xg_res, "rmse")`.

### How well did the xgboost model do?

```{r}
collect_predictions(final_xg_res) %>% 
    ggplot(aes(imdb, .pred)) +
    geom_point(alpha = 0.6, color = "midnightblue") +
    geom_abline(lty = 2, color = "red") +
    coord_obs_pred() +
    labs(x = "imdb score (observed)",
         y = "imdb score (predicted)")
```

Let's try a simpler model

```{r}

scoobydoo_rec_basic <- recipe(imdb ~ ., data = scoobydoo_train) %>% 
    update_role(index, new_role = "id") %>% 
    step_mutate(date_aired = 10 * (year(date_aired) %/% 10),
                monster_real = factor(monster_real),
                snack_shaggy = factor(snack_shaggy)) %>% 
    step_other(all_nominal_predictors(), threshold = 0.1) %>% 
    step_impute_median(all_numeric_predictors(), all_outcomes()) %>% 
    step_impute_mode(all_nominal_predictors()) %>% 
    step_log(engagement, base = 10, offset = 1) %>% 
    step_unknown(snack_shaggy, new_level = "unknown") %>%
    step_unknown(monster_real, new_level = "unknown") %>% # since snack_shaggy and monster_real still have missing values even after imputation, I set the missing value as unknown.
    step_dummy(all_nominal_predictors())

scoobydoo_rec_basic %>% 
    prep() %>% 
    juice()
```


```{r}

scoobydoo_xg_wf_basic <- workflow() %>% 
    add_recipe(scoobydoo_rec_basic) %>% 
    add_model(scoobydoo_spec_xg)
```

```{r}
set.seed(2021)
scoobydoo_xg_res_basic <- scoobydoo_xg_wf_basic %>% 
    tune_grid(scoobydoo_fold,
              grid = crossing(mtry = seq(10, 35, 5),
                              trees = seq(300, 1000, 100),
                              tree_depth = seq(5, 15, 5),
                              learn_rate = c(0.01, 0.02)),
              control = control_grid(save_pred = TRUE,
                                     save_workflow = TRUE))

autoplot(scoobydoo_xg_res_basic)

scoobydoo_xg_res_basic %>% 
    collect_metrics() %>% 
    arrange(mean)
```

### Explore results

```{r}
show_best(scoobydoo_xg_res_basic, "rmse")
show_best(scoobydoo_xg_res_basic, "rsq")
```

* The best rmse for training data tuned by xgboost model is `show_best(scoobydoo_xg_res_basic, "rmse")`.

### Finalize xgboost workflow
```{r}
final_xg_wf_basic <- finalize_workflow(scoobydoo_xg_wf_basic,
                      select_best(scoobydoo_xg_res_basic, "rmse"))
```


### Important variables
```{r}
library(vip)
final_xg_wf_basic %>% 
    fit(data = scoobydoo_train) %>% 
    pull_workflow_fit() %>% 
    vip(geom = "col")
```

### Fitting on test data

```{r}
final_xg_res_basic <- last_fit(final_xg_wf_basic, scoobydoo_split)
final_xg_res_basic %>% 
    collect_metrics()
```

* The best rmse for testing data tuned by xgboost basic model is `show_best(final_xg_res_basic, "rmse")`.

### How well did the xgboost basic model do?

```{r}
collect_predictions(final_xg_res_basic) %>% 
    ggplot(aes(imdb, .pred)) +
    geom_point(alpha = 0.6, color = "midnightblue") +
    geom_abline(lty = 2, color = "red") +
    coord_obs_pred() +
    labs(x = "imdb score (observed)",
         y = "imdb score (predicted)")
```

xgboost basic model performs better.