---
title: "TidyTemplate"
date: 2021-07-29
output: 
    html_document:
        toc: true
        toc_float: true
        theme: flatly
        highlight: pygments
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-07-27")
olympics <- tt$olympics
regions <- tt$regions
olympic_data <- olympics %>% 
    left_join(regions, by = c("noc" = "NOC"))
```

## EDA

```{r}
olympic_data %>% 
    skimr::skim()
```

### Total medals awarded

```{r}
olympic_data %>% 
    filter(!is.na(medal)) %>% 
    count(medal, sort = TRUE)
```

* Interestingly, there are more Gold medals awarded than Bronze and Silver. Possibly, people got the same results and tied, so they got the same medals.

#### How about total awarded medals breaking down by gender

```{r}
olympic_data %>% 
    filter(!is.na(medal)) %>% 
    count(medal, sex) %>% 
    mutate(medal = fct_relevel(medal, "Silver", after = 1)) %>% 
    ggplot(aes(n, medal, fill = sex)) +
    geom_col(position = "dodge", width = 0.3, alpha = 0.7) +
    geom_text(aes(label = n), vjust = 1.2, hjust = 1) +
    scale_fill_manual(values = c("#ED008C", "#03ACED")) +
    labs(x = "# of medals",
         y = NULL,
         fill = "Gender",
         title = "# of medals awarded in Olympics by type and gender") +
    ggthemes::theme_clean() +
    theme(legend.position = "top")
```

* Men were awarded much more medals than women. Probably, it is because men participated in the Olympics much earlier than women. This assumption can be tested below

### Did men participated in the Olympics much earlier than women?
```{r}
olympic_data %>% 
    filter(!is.na(medal)) %>% 
    group_by(sex) %>% 
    slice_min(year, n = 1, with_ties = FALSE)
```

It turns out that the assumption is not true since the first participation for male athletes was 1896 while it was 1900 for females. Could it be the issue of the variety of sports for female athletes are fewer than their male counterparts? 

### Type of sports for men and women in Olympics

```{r}
library(tidytext)
library(plotly)
p <- olympic_data %>% 
    count(sport, sex) %>% 
    mutate(sport = reorder_within(sport, n, sex)) %>%
    mutate(sex = fct_recode(sex, "Female" = "F", "Male" = "M")) %>% 
    ggplot(aes(n, sport, fill = sex)) +
    geom_col() +
    facet_wrap(~sex) +
    scale_y_reordered() +
    scale_fill_manual(values = c("#ED008C", "#03ACED")) +
    labs(x = "# of medals",
         y = NULL,
         fill = "Gender",
         title = "# of medals awarded in Olympics by sport and gender") +
    theme(legend.position = "none")

ggplotly(p, tooltip = c("y", "x"))  
```

It is clear that male athletes have more types of sports to participate in comparing to female. Another factor could be the number of total athletes as well.

### Total number of athletes by gender

```{r}
olympic_data %>% 
    count(sex, sort = TRUE)
```

Wow, the answer is that male athletes is `196594/74522` times more than female athletes.

### Medals by team

```{r}
olympic_data %>% 
  filter(!is.na(medal)) %>% 
  mutate(medal = fct_relevel(medal, "Bronze", after = 2)) %>% 
  group_by(team, medal) %>% 
  summarize(total_medal = n()) %>%
  ungroup() %>% 
  group_by(medal) %>% 
  slice_max(total_medal, n = 20) %>% 
  mutate(team = reorder_within(team, total_medal, medal)) %>%
  ggplot(aes(total_medal, team, fill = medal)) +
  geom_col() +
  geom_text(aes(label = total_medal), hjust = 1) +
  facet_wrap(~medal, scales = "free") +
  scale_fill_manual(values = c("#DFB90D", "#9AA0A6", "#AE7058")) +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  labs(x = "# of medals",
       y = NULL,
       fill = NULL,
       title = "Total medals by team") +
  ggthemes::theme_fivethirtyeight() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank())
```

### Medals by team over time

```{r}
library(gganimate)
medal <- olympic_data %>% 
  filter(!is.na(medal)) %>% 
  mutate(medal = fct_relevel(medal, "Bronze", after = 2)) %>% 
  group_by(year, team, medal) %>% 
  summarize(total_medal = n()) %>% 
  ungroup() %>% 
  filter(team %in% c("United States", "Soviet Union", "Germany", "Italy",
                     "Great Britain", "France", "Canada", "China", "Japan",
                     "Australia")) %>% 
  ggplot(aes(total_medal, team, group = team, fill = medal)) +
  geom_col() +
  scale_fill_manual(values = c("#DFB90D", "#9AA0A6", "#AE7058")) +
  transition_manual(frames = year) +
  labs(x = "# of medals",
       y = NULL,
       fill = "Medals",
       title = "Total medals by team over time",
       subtitle = "Year: {current_frame}") +
  ggthemes::theme_clean() +
  theme(legend.position = "top",
        panel.grid.major.y = element_blank())

animate(medal, fps = 1, 
        renderer = gifski_renderer(loop = TRUE)) 
  
```

## Animation inspired by Katie Press

```{r}
animation_df <- olympic_data %>% 
  count(year, season, region, event, medal) %>% 
  count(year, season, region, medal) %>% 
  filter(!is.na(medal)) %>% 
  rename("num_medal" = n) # getting the number of medal by country, year, and medal type

animation_df <- animation_df %>% 
  group_by(year, season, region) %>% 
  mutate(total_medal = sum (num_medal)) %>%  # getting the total number of medal by country and year
  ungroup()

total_athletes <- olympic_data %>% 
  count(year, season, region, sex, age, name) %>% 
  count(year, season, region) %>% 
  rename("total_athletes" = n) # count total athletes for each country by year

animation_df <- animation_df %>% 
  left_join(total_athletes)
```

### Find the top performing countries in summer olympics

```{r}
top_10_countries <- animation_df %>% 
  filter(season == "Summer", year > 1976) %>% 
  distinct(year, region, total_medal) %>% 
  group_by(region) %>% 
  summarize(avg_medals = round(mean(total_medal), 2)) %>% 
  arrange(desc(avg_medals)) %>% 
  slice(1:10) %>% 
  pull(region)
```

### Generate the rank for top ten countries
```{r}
medals_df <- animation_df %>% 
  filter(season == "Summer", year > 1979, region %in% top_10_countries) %>% 
  distinct(year, region, total_medal) %>% 
  arrange(year, desc(total_medal)) %>% 
  group_by(year) %>% 
  mutate(rank = seq_along(total_medal))

medals_df %>% 
  print(n = Inf)
# The problem is that some countries won the same number of medals such as 
# Australia, Italy, and Japn all won 14 medals in 1988
```

```{r}
# Use rank from last Olympic to solve the tied medals issue
medals_df_lagged <-  medals_df %>% 
  arrange(region, year) %>% 
  group_by(region) %>% 
  mutate(prev_rank = lag(rank)) %>% 
  arrange(year, desc(total_medal), prev_rank) %>% 
  group_by(year) %>% 
  mutate(final_rank = seq_along(region)) %>% 
  ungroup()

medals_df_lagged %>% print(n = Inf)
```

```{r}
library(ggflags)
library(countrycode)
final_df <- medals_df_lagged %>% 
  filter(year >= 1988) %>% 
  mutate(country_code = str_to_lower(countrycode(region, origin = "country.name",
                                    destination = "iso2c")))
medal_line <- final_df %>% 
  ggplot(aes(x = year, y = final_rank, color = region)) +
  geom_point(aes(group = seq_along(final_rank), color = region), size = 4) +
  geom_line(aes(color = region), alpha = 0.6, size = 2) +
  scale_x_continuous(breaks = seq(1980, 2016, 4))+
  scale_y_reverse(breaks = seq(1, 10, 1))+ 
  geom_text(aes(label = region), hjust = -0.5) +
  geom_flag(data = final_df,
            country = final_df$country_code,
            size = 8) +
  labs(x = NULL,
       y = "Rank",
       color = NULL,
       title = "Olympic medals by country and year") +
  transition_reveal(year) +
  ggthemes::theme_hc() +
  theme(legend.position = "none")

animate(medal_line, fps = 6,
        renderer = gifski_renderer(loop = TRUE)) 
```

```{r}
anim_save("2021_07_27_tidy_tuesday_olympics.gif")
```

