---
title: "TidyTemplate"
date: 2021-06-15
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(tidytuesdayR)
library(DataExplorer)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-06-15")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tweets <- tt$tweets
```

## EDA

```{r}
tweets %>% 
    skimr::skim()
```

```{r}
tweets %>% 
    plot_intro()
```

```{r}
tweets %>% 
    plot_missing()
```

## Which user tweeted the most?

```{r}
tweets %>% 
    count(username, sort = TRUE) %>% 
    slice_max(n, n = 10) %>% 
    mutate(username = fct_reorder(username, n)) %>%
    ggplot(aes(n, username)) +
    geom_col() +
    geom_text(aes(label = n), hjust = -0.1) +
    labs(x = "# of tweets",
         y = "",
         title = "Top 10 most tweeted users")
```

## Top users' stats by arranged by median followers
```{r}
tweets %>% 
    group_by(username) %>% 
    summarize(median_retweet = median(retweet_count),
              median_like = median(like_count),
              median_quote = median(quote_count),
              median_followers = median(followers)) %>% 
    arrange(desc(median_followers))
```


## Top users' stats by arranged by median retweet
```{r}
tweets %>% 
    group_by(username) %>% 
    summarize(median_retweet = median(retweet_count),
              median_like = median(like_count),
              median_quote = median(quote_count),
              median_followers = median(followers)) %>% 
    arrange(desc(median_retweet))
```

## Top users' stats by arranged by median like
```{r}
tweets %>% 
    group_by(username) %>% 
    summarize(median_retweet = median(retweet_count),
              median_like = median(like_count),
              median_quote = median(quote_count),
              median_followers = median(followers)) %>% 
    arrange(desc(median_like))
```

## Top users' stats by arranged by median quote
```{r}
tweets %>% 
    group_by(username) %>% 
    summarize(median_retweet = median(retweet_count),
              median_like = median(like_count),
              median_quote = median(quote_count),
              median_followers = median(followers)) %>% 
    arrange(desc(median_quote))
```

- User with the highest median retweet count also has the highest median like and median quote.

## Who and which tweet has the highest retweet

```{r}
tweets %>% 
    slice_max(retweet_count, n = 1) %>% 
    select(retweet_count, content, username) 
```

## Who and which tweet has the highest like

```{r}
tweets %>% 
    slice_max(like_count, n = 1) %>% 
    select(like_count, content, username) 
```

## Who and which tweet has the highest quote

```{r}
tweets %>% 
    slice_max(quote_count, n = 1) %>% 
    select(quote_count, content, username)
```

- User **CharlieEatonPhD** has the highest retweet, like, and quote count.

## Question: What factors influence the retweet count?
### Does the user matter?

```{r}
tweets %>% 
    filter(!is.na(username)) %>% 
    mutate(username = fct_lump(username, n = 9),
           username = fct_reorder(username, retweet_count)) %>% 
    ggplot(aes(retweet_count + 1, username)) +
    geom_boxplot(aes(fill = username), alpha = 0.5)+
    scale_x_log10() +
    labs(x = "Retweet count (logged)",
         y = "",
         title = "Retweet count by most prevalent users") +
    theme(legend.position = "none")
```
- It seems that retweet count is associated with different users

### How about like and quote counts as well as followers and locations in terms of longitude and latitude

```{r}
tweets %>% 
    na.omit() %>% 
    plot_correlation(type = "continuous")
```
- It is clear that like and quote counts are extremely strong predictors of retweet count, while the variable `followers` has a positively weak correlation with retweet count. The locations seem not to matter much given the correlations for longitude and latitude with retweet count are very weak and negative.

### Visualize the relationship between retweet count and location

```{r}
library(leaflet)

pal <- colorFactor(palette = "RdBu",
                    domain = tweets$retweet_dummy)
tweets_map <- tweets %>% 
    mutate(popup_label = paste(paste0("<b>Location: ", "</b>",
                                      location),
                               paste0("<b>Retweet Count: ", "</b>",
                                      retweet_count),
                               sep = "<br/>"),
           retweet_dummy = if_else(retweet_count > 10, "High", "Low")) %>% 
    na.omit()

leaflet() %>% 
    addProviderTiles("Stamen.TonerLite") %>% 
    addCircleMarkers(data = tweets_map,
                     radius = 10,
                     stroke = FALSE,
                     fillOpacity = 0.7,
                     popup = ~popup_label,
                     fillColor = ~pal(retweet_dummy)) %>% 
    addLegend(data = tweets_map,
              position = "topright",
              values = ~retweet_dummy,
              title = "Retweet Count",
              pal = pal)
```

### Does datetime matter?

```{r}
# Look at the retweet count by each day
library(lubridate)
tweets %>% 
    mutate(date = date(datetime)) %>% 
    group_by(date) %>% 
    summarize(retweet_day = median(retweet_count)) %>% 
    ungroup() %>% 
    ggplot(aes(date, retweet_day)) +
    geom_line() +
    geom_smooth(method = "lm") +
    labs(x = "Date",
         y = "# of retweets (median)",
         title = "Median # of retweets by day",
         subtitle = "Feb, 2021 - May, 2021")
```

- It does not seem to have a linear relationship between the time of the tweets and retweet count.

```{r}
# Look at the retweet count by each week
tweets %>% 
    mutate(week = week(datetime)) %>% 
    group_by(week) %>% 
    summarize(retweet_week = median(retweet_count)) %>% 
    ungroup() %>% 
    ggplot(aes(week, retweet_week)) +
    geom_line() +
    geom_smooth(method = "lm") +
    labs(x = "Week",
         y = "# of retweets (median)",
         title = "Median # of retweets by week",
         subtitle = "Feb, 2021 - May, 2021")
```

- A negative association seems to exist when we organize the `datetime` by week. The median of retweet dropped in recently weeks from about 1.3 per week to less than 1 per week.


## Model
### Train, test, and validation sets

```{r}
library(tidymodels)
set.seed(121)
tweets_split <- initial_split(tweets)
tweets_train <- training(tweets_split)
tweets_test <- testing(tweets_split)
tweets_fold <- vfold_cv(tweets_train, v = 10)
```

### Feature engineering

```{r}
library(textrecipes)
tweets_rec <- tweets_train %>% 
    recipe(retweet_count ~ like_count + quote_count + followers + content
           + username) %>% 
    step_log(all_numeric_predictors(), all_outcomes(), base = 10,
             offset = 1) %>% 
    step_other(username, threshold = 5) %>% 
    step_dummy(username, one_hot = TRUE) %>% 
    step_tokenize(content, token = "tweets") %>% 
    step_stopwords(content, stopword_source = "stopwords-iso") %>% 
    step_tokenfilter(content, max_tokens = 10) %>% 
    step_tfidf(content) 

tweets_rec %>% 
    prep() %>% 
    juice() %>% View()
```

