---
title: "2021_07_20_tidy_tuesday"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(lubridate)
theme_set(theme_light())
```


## Load data

```{r}
drought <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv') %>% 
    mutate(map_date = parse_date(as.character(map_date), format = "%Y%m%d"),
           drought_lvl = fct_recode(drought_lvl,
                                    "No drought" = "None",
                                    "Abnormally dry" = "D0",
                                    "Moderate drought" = "D1",
                                    "Severe drought" = "D2",
                                    "Extreme drought" = "D3",
                                    "Exceptional drought" = "D4"
                                    ),
           drought_lvl = fct_relevel(drought_lvl, c("No drought",
                                                    "Abnormally dry",
                                                    "Moderate drought",
                                                    "Severe drought",
                                                    "Extreme drought",
                                                    "Exceptional drought")))
```

## EDA

```{r}
drought %>% 
    skimr::skim()
```

### Minimum and Maximum recorded date for each state

```{r}
drought %>% 
    group_by(state_abb) %>% 
    summarize(min_date = min(map_date),
              max_date = max(map_date),
              mean_date = mean(map_date),
              median_date = median(map_date)) %>% 
    print(n = Inf)
```

* Each state has the same amount of recorded date

### When was the most prevalent drought (Abnormally dry) in each state?

```{r}
library(geofacet)
drought %>% 
    filter(drought_lvl == "Abnormally dry") %>% 
    group_by(state_abb) %>% 
    slice_max(area_pct, n = 1) %>% 
    ggplot(aes(map_date, area_pct/100)) +
    geom_point() +
    scale_y_continuous(labels = percent) +
    facet_geo(~state_abb) +
    labs(x = "Drought date",
         y = "% of state in drought",
         title = "% of state in drought (abnormally dry) over time") +
    theme(axis.text.x = element_text(angle = 30))
```

* D.C., New Jersey, and Rhode Island have the most prevalent drought (abnormally dry) affecting almost 100% of the state (or special district) on multiple occasions.

### The drought (abnormally dry) hits the most population in each state

```{r}
drought %>% 
    filter(drought_lvl == "Abnormally dry") %>% 
    group_by(state_abb) %>% 
    slice_max(pop_pct, n = 1) %>% 
    ggplot(aes(map_date, pop_pct/100)) +
    geom_point() +
    scale_y_continuous(labels = percent) +
    facet_geo(~state_abb) +
    labs(x = "Drought date",
         y = "% of population in drought",
         title = "% of population in drought (abnormally dry) over time in each state") +
    theme(axis.text.x = element_text(angle = 30))
```

* Similar trend was identified.


### All drought conditions over time

```{r}
drought %>% 
    ggplot(aes(map_date, area_pct/100, fill = drought_lvl)) +
    geom_area() +
    facet_geo(~state_abb, label = "name") +
    scale_fill_viridis_d(option = "A") +
    scale_y_continuous(labels = percent) +
    labs(title = "Droughts in the U.S. (2001 - 2021)",
         subtitle = "% of area in drought",
         fill = "",
         x = "",
         y = "") +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, face = "bold", size = 24),
          plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 16),
          axis.text.x = element_text(angle = 30))
```


### Using plotly to visualize only extreme drought

```{r}
library(plotly)
drought %>%
    filter(drought_lvl == "Extreme drought") %>% 
    group_by(state_abb, year(map_date), month(map_date, label = TRUE)) %>% 
    summarize(`Average area impacted` = mean(area_pct)) %>% 
    ungroup() %>% 
    mutate(year = `year(map_date)`, month = `month(map_date, label = TRUE)`) %>% 
    unite("time", year:month,
                        sep = ", ") %>% 
    plot_ly(type = "choropleth",
            locationmode = "USA-states",
            locations = ~state_abb,
            z = ~`Average area impacted`,
            frame = ~time,
            color = ~`Average area impacted`,
            zauto = FALSE,
            zmin = 0,
            zmax = 100) %>% 
    layout(
    title = "Extreme droughts in the US",
    geo = list(
    scope = 'usa'),
    showlegend = FALSE) %>% 
    colorbar(title = "Avg. % of area impacted")
```

