---
title: "TidyTemplate"
date: 2021-05-11
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-05-11")
broadband_county <- tt$broadband %>% 
    janitor::clean_names() %>% 
    mutate(broadband_availability_per_fcc = as.numeric(broadband_availability_per_fcc),
           broadband_usage = as.numeric(broadband_usage))
broadband_zip <- tt$broadband_zip %>% 
    janitor::clean_names()
```

## Counties with highest broadband access in each state
```{r}
library(tidytext)
broadband_county %>% 
    group_by(st) %>% 
    slice_max(broadband_availability_per_fcc, n = 3, with_ties = FALSE) %>% 
    mutate(county_name = reorder_within(county_name,
                                        broadband_availability_per_fcc,
                                        st)) %>%
    ungroup() %>% 
    ggplot(aes(broadband_availability_per_fcc, county_name,
               fill = county_name)) +
    geom_col() +
    facet_wrap(~st, scales = "free", ncol = 6) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per county with access to broadband",
         y = "",
         title = "Counties with highest broadband accessibility in each state") +
    theme(legend.position = "none")
```


## Counties with lowest broadband access in each state

```{r}
broadband_county %>% 
    group_by(st) %>% 
    slice_min(broadband_availability_per_fcc, n = 3, with_ties = FALSE) %>% 
    mutate(county_name = reorder_within(county_name,
                                        broadband_availability_per_fcc,
                                        st)) %>% 
    ungroup() %>% 
    ggplot(aes(broadband_availability_per_fcc, county_name,
               fill = county_name)) +
    geom_col() +
    facet_wrap(~st, scales = "free", ncol = 6) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per county with access to broadband",
         y = "",
         title = "Counties with Lowest broadband accessibility in each state") +
    theme(legend.position = "none")
```

## Counties with highest broadband usage in each state

```{r}
broadband_county %>% 
    group_by(st) %>% 
    slice_max(broadband_usage, n = 3, with_ties = FALSE) %>% 
    mutate(county_name = reorder_within(county_name,
                                        broadband_usage,
                                        st)) %>% 
    ungroup() %>% 
    ggplot(aes(broadband_usage, county_name,
               fill = county_name)) +
    geom_col() +
    facet_wrap(~st, scales = "free", ncol = 6) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per county use broadband",
         y = "",
         title = "Counties with highest broadband usage in each state") +
    theme(legend.position = "none")
```

## Counties with lowest broadband usage in each state

```{r}
broadband_county %>% 
    group_by(st) %>% 
    slice_min(broadband_usage, n = 3, with_ties = FALSE) %>% 
    mutate(county_name = reorder_within(county_name,
                                        broadband_usage,
                                        st)) %>% 
    ungroup() %>% 
    ggplot(aes(broadband_usage, county_name,
               fill = county_name)) +
    geom_col() +
    facet_wrap(~st, scales = "free", ncol = 6) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per county use broadband",
         y = "",
         title = "Counties with lowest broadband usage in each state") +
    theme(legend.position = "none")
```

## County map

```{r}
library(tidygeocoder)
library(leaflet)
library(leaflegend)
# broadband_county_geocoded <- broadband_county %>% 
#     geocode(county = county_name,
#             state = st,
#             method = "osm",
#             verbose = TRUE)
#write_csv(broadband_county_geocoded, "broadband_county_geocoded.csv")

broadband_county_geocoded <- read_csv("broadband_county_geocoded.csv")

broadband_county_geocoded <- broadband_county_geocoded %>% 
    mutate(popup_label = paste(paste0("<b>County: ", "</b>",
                                      county_name, ", ", st),
                               paste0("<b>Broadband availability: ", "</b>",
                                      broadband_availability_per_fcc*100, "%"),
                               paste0("<b>Broadband usage: ", "</b>",
                                      broadband_usage*100, "%"),
                               sep = "<br/>")) %>% 
    filter(!is.na(broadband_availability_per_fcc),
           !is.na(broadband_usage))
           
pal <- colorNumeric(palette = "RdYlBu",
                    domain = broadband_county_geocoded$broadband_usage*100)
sizes <- sizeNumeric(values = broadband_county_geocoded$broadband_availability_per_fcc*100,
                     baseSize = 0)

leaflet() %>% 
    addProviderTiles("Stamen.TonerLite") %>% 
    addCircleMarkers(data = broadband_county_geocoded,
                     radius = ~ broadband_availability_per_fcc*10,
                     stroke = FALSE,
                     fillOpacity = 0.7,
                     popup = ~popup_label,
                     fillColor = ~pal(broadband_usage*100)) %>%
    addLegendSize(values = broadband_county_geocoded$broadband_availability_per_fcc*100,
                  position = "topright",
                  pal = pal,
                  title = "Broadband Availability",
                  shape = c("circle"),
                  orientation = c('vertical', 'horizontal'),
                  opacity = 0.7,
                  breaks = 6) %>% 
       addLegend(data = broadband_county_geocoded,
              position = "topright",
              values = ~broadband_usage*100,
              title = "Broadband Usage",
              pal = pal)
```

### County choropleth map

```{r}
library(tigris)
library(sf)
counties_sf <- counties(state = unique(broadband_county$st),
                        cb = TRUE,
                        resolution = "500k",
                        year = 2019) 

# Add leading zeros to county_id to join with GEOID in counties_sf

broadband_county <- broadband_county %>% 
     mutate(county_id = sprintf("%05d", county_id)) #c-style string formatting

broadband_county_sf <- counties_sf %>% 
    right_join(broadband_county, by = c("GEOID" = "county_id"))



mytext <- paste(paste0("<b>County: </b>", broadband_county_sf$county_name, ", ",
                broadband_county_sf$st),
                paste0("<b>Broadband Usage: </b>",
                broadband_county_sf$broadband_usage*100, "%"),
                sep = "<br/>") %>% 
    lapply(htmltools::HTML)

pal <- colorBin(palette = "RdYlBu",
                    domain = broadband_county_sf$broadband_usage*100,
                 bins = 5)

leaflet() %>%
    addProviderTiles("Stamen.TonerLite") %>%
    addPolygons(data = broadband_county_sf,
                stroke = TRUE,
                fillColor = ~pal(broadband_usage*100),
                fillOpacity = 0.9,
                color = "white",
                weight = 0.3,
                label = mytext,
                labelOptions = labelOptions(
                    style = list("font-weight" = "normal",
                                 padding = "3px 8px"),
                    textsize = "13px",
                    direction = "auto" 
                )) %>% 
    addLegend(pal = pal,
              values = broadband_county_sf$broadband_usage,
              opacity = 0.9,
              title = paste("Percentage of people using the internet", 
              "at 25Mbps or above per county",
              sep = "<br/>"),
              labFormat = labelFormat(suffix = "%"),
              position = "bottomleft")
    
```

### mapping with tmap (static)

```{r}
library(tmap)
us_states_map <- tm_shape(broadband_county_sf %>% 
             filter(!st %in% c("AK", "HI")), projection = 2163) +
    tm_polygons(col = "broadband_usage",
                palette = "RdYlBu",
                title = "% broadband usage")  +
    tm_layout(frame = FALSE,
              legend.position = c("right", "top"))

hawaii_map <- tm_shape(broadband_county_sf %>% 
             filter(st %in% c("HI"))) +
    tm_polygons(col = "broadband_usage",
                palette = "RdYlBu")  +
    tm_layout(frame = FALSE,
              legend.show = FALSE)

alaska_map <- tm_shape(broadband_county_sf %>% 
             filter(st %in% c("AK"))) +
    tm_polygons(col = "broadband_usage",
                palette = "RdYlBu")  +
    tm_layout(frame = FALSE,
              legend.show = FALSE)

us_states_map
print(hawaii_map, vp = grid::viewport(0.35, 0.1, width = 0.2, height = 0.1))
print(alaska_map, vp = grid::viewport(0.15, 0.15, width = 0.3, height = 0.3))

```

### mapping with tmap (interactive)

```{r}
library(tmap)
tmap_mode("view")
tm_shape(broadband_county_sf) +
    tm_polygons(col = "broadband_usage",
                palette = "RdYlBu",
                title = "% broadband usage",
                popup.vars = c("county_name", "st", "broadband_usage")) +
    tm_layout(frame = FALSE,
              legend.position = c("right", "top"),
              legend.format = list(
                  suffix = "%")) +
    tm_basemap(server = "Stamen.TonerLite")

```

### Show broadband availability and usage seperately
```{r}
pal1 <- colorNumeric(palette = "YlGnBu",
                    domain = broadband_county_geocoded$broadband_availability_per_fcc*100)
leaflet() %>% 
    addProviderTiles('CartoDB', group = 'CartoDB') %>% 
    addProviderTiles('Esri.WorldGrayCanvas', group = 'Esri.WorldGrayCanvas') %>%       addCircleMarkers(data = broadband_county_geocoded,
                     radius = 10,
                     stroke = FALSE,
                     fillOpacity = 0.7,
                     popup = ~popup_label,
                     fillColor = ~pal1(broadband_availability_per_fcc*100),
                     group = "broadband_availability") %>% 
    addCircleMarkers(data = broadband_county_geocoded,
                     radius = 10,
                     stroke = FALSE,
                     fillOpacity = 0.7,
                     popup = ~popup_label,
                     fillColor = ~pal(broadband_usage*100),
                     group = "broadband_usage") %>%
    addLayersControl(baseGroups = c('CartoDB', 'Esri.WorldGrayCanvas'),
                     overlayGroups = c("broadband_availability",
                                       "broadband_usage"))
```

## Broadband zipcode

```{r}
library(zipcodeR)
# zip_geocode <- apply(broadband_zip %>% select(postal_code), 2, FUN = geocode_zip)
zip_geocode <- map_df(broadband_zip %>% select(postal_code), geocode_zip) %>% 
    mutate(zipcode = as.numeric(zipcode))
broadband_zip_geocoded <- broadband_zip %>% 
    left_join(zip_geocode, by = c("postal_code" = "zipcode"))
```




## Zipcode with highest broadband usage in each state

```{r}
broadband_zip_geocoded %>% 
    group_by(st) %>% 
    slice_max(broadband_usage, n = 3, with_ties = FALSE) %>% 
    mutate(postal_code = reorder_within(postal_code,
                                        broadband_usage,
                                        st)) %>% 
    ungroup() %>% 
    ggplot(aes(broadband_usage, postal_code,
               fill = postal_code)) +
    geom_col() +
    geom_text(aes(label = county_name), size = 3, hjust = 1.5) +
    facet_wrap(~st, scales = "free") +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per zipcode use broadband",
         y = "",
         title = "Zipcode with highest broadband usage in each state") +
    theme(legend.position = "none")
```


## Zipcode with lowest broadband usage in each state

```{r}
broadband_zip_geocoded %>% 
    group_by(st) %>% 
    slice_min(broadband_usage, n = 3, with_ties = FALSE) %>%
    mutate(postal_code = reorder_within(postal_code,
                                        broadband_usage,
                                        st)) %>% 
    ungroup() %>% 
    ggplot(aes(broadband_usage, postal_code,
               fill = postal_code)) +
    geom_col() +
    geom_text(aes(label = county_name), size = 3, hjust = 1) +
    facet_wrap(~st, scales = "free", ncol = 6) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    labs(x = "Percentage of people per zipcode use broadband",
         y = "",
         title = "Zipcode with lowest broadband usage in each state") +
    theme(legend.position = "none")
```

## Zipcode with highest broadband usage in each county
```{r}
broadband_zip_geocoded %>%
    filter(!st %in% c("AK", "HI")) %>% 
    group_by(st, county_name) %>% 
    slice_max(broadband_usage, n = 1, with_ties = FALSE) %>% 
    ungroup() %>% 
    ggplot(aes(lng, lat, color = broadband_usage)) +
    geom_point() +
    borders(database = "county") +
    scale_color_gradient2(name = "Broadband Usage",
                          low = "blue",
                          mid = "white",
                          high = "red",
                          midpoint = 0.5,
                          space = "Lab",
                          labels = percent) +
    labs(title = "Zipcode with highest broadband usage in each county",
         x = "Longitude",
         y = "Latitude") +
    ggthemes::theme_fivethirtyeight()
    
```

## Zipcode with lowest broadband usage in each county

```{r}
broadband_zip_geocoded %>%
    filter(!st %in% c("AK", "HI")) %>% 
    group_by(st, county_name) %>% 
    slice_min(broadband_usage, n = 1, with_ties = FALSE) %>% View()
    ungroup() %>% 
    ggplot(aes(lng, lat, color = broadband_usage)) +
    geom_point() +
    borders(database = "county") +
    scale_color_gradient2(name = "Broadband Usage",
                          low = "blue",
                          mid = "white",
                          high = "red",
                          midpoint = 0.5,
                          space = "Lab",
                          labels = percent) +
    labs(title = "Zipcode with lowest broadband usage in each county",
         x = "Longitude",
         y = "Latitude") +
    ggthemes::theme_fivethirtyeight()
```


## Map

```{r}
broadband_zip_geocoded <- broadband_zip_geocoded %>% 
    mutate(popup_label = paste(paste0("<b>Zipcode: ", "</b>",
                                      postal_code),
                               paste0("<b>County: ", "</b>",
                                      county_name, ", ", st),
                               paste0("<b>Broadband usage: ", "</b>",
                                      broadband_usage*100, "%"),
                               sep = "<br/>")) %>% 
    filter(!is.na(broadband_usage))

pal <- colorNumeric(palette = "RdYlBu",
                    domain = broadband_zip_geocoded$broadband_usage*100)

leaflet() %>% 
    addProviderTiles("Stamen.TonerLite") %>% 
    addCircleMarkers(data = broadband_zip_geocoded,
                     radius = 4,
                     stroke = FALSE,
                     fillOpacity = 0.5,
                     popup = ~popup_label,
                     fillColor = ~pal(broadband_usage*100)) %>%
       addLegend(data = broadband_zip_geocoded,
              position = "topright",
              values = ~broadband_usage*100,
              title = "Broadband Usage",
              pal = pal)
```

