---
title: "TidyTemplate"
date: 2021-08-07
output: 
    html_document:
        toc: true
        toc_float: true
        theme: flatly
        highlight: pygments
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-08-03")
paralympics <- tt$athletes
```

## Data Cleaning

```{r}
paralympics %>% 
    count(abb, sort = TRUE) %>% 
    print(n = Inf)
# A lof of NAs in country but not for abb so we use countrycode package to find country names

library(countrycode)
para_cleaned <- paralympics %>% 
    mutate(country_fixed = countrycode(abb, origin = "ioc", destination = "country.name"),
           country_iso2c = str_to_lower(countrycode(abb, origin = "ioc",
                                                    destination = "iso2c")),
           country_fixed = case_when(abb == "BIR" ~ "Myanmar",
                                     abb == "EUN" ~ "United Team",
                                     abb == "FRG" ~ "West Germany",
                                     abb == "GDR" ~ "East Germany",
                                     abb == "IPP" ~ "Independent Paralympic Participants",
                                     abb == "SCG" ~ "Serbia and Montenegro",
                                     abb == "TCH" ~ "Czechoslovakia",
                                     abb == "URS" ~ "Soviet Union",
                                     abb == "YUG" ~ "Yugoslavia",
                                     TRUE ~ country_fixed),
           country_iso2c = str_to_lower(countrycode(country_fixed, origin = "country.name",
                                                    destination = "iso2c")))
# Czechoslovakia, East Germany, Independent Paralympic Participants, Serbia and Montenegro, United Team, Yugoslavia cannot be matched with 2 character iso code
```

## Animation

```{r}
animation_df <- para_cleaned %>% 
    count(year, country_fixed, event, medal) %>% 
    count(year, country_fixed, medal) %>% 
    rename("num_medals" = n)

animation_df <- animation_df %>% 
    group_by(year, country_fixed) %>% 
    mutate(total_medals = sum(num_medals)) %>% 
    ungroup()
```

### Find the top 10 performing countries

```{r}
(top_10_countries <- animation_df %>% 
  distinct(year, country_fixed, total_medals) %>% 
  group_by(country_fixed) %>% 
  summarize(avg_medals = round(mean(total_medals), 2)) %>% 
  arrange(desc(avg_medals)) %>% 
  slice(1:11) %>% 
  pull(country_fixed))
# used top 11 because West Germany and east Germany united after 1990
```

### Generate rank for top countries

```{r}
medals_df <- animation_df %>% 
  filter(country_fixed %in% top_10_countries) %>% 
  distinct(year, country_fixed, total_medals) %>% 
  arrange(year, desc(total_medals)) %>% 
  group_by(year) %>% 
  mutate(rank = seq_along(total_medals))

medals_df %>% 
  print(n = Inf)
# The problem is that some countries won the same number of medals such as 
# Canada, and Spain all won 14 medals in 1988
```


```{r}
# use rank from last Paralympic to solve the tied medals issue
medals_df_lagged <-  medals_df %>% 
  arrange(country_fixed, year) %>% 
  group_by(country_fixed) %>% 
  mutate(prev_rank = lag(rank)) %>% 
  arrange(year, desc(total_medals), prev_rank) %>% 
  group_by(year) %>% 
  mutate(final_rank = seq_along(country_fixed)) %>% 
  ungroup()

medals_df_lagged %>% print(n = Inf)
```


```{r}
library(ggflags)
library(gganimate)

final_df <- medals_df_lagged %>% 
  filter(year >= 1996) %>% 
  mutate(country_code = str_to_lower(countrycode(country_fixed, origin = "country.name",
                                    destination = "iso2c")))
medal_line <- final_df %>% 
  ggplot(aes(x = year, y = final_rank, color = country_fixed)) +
  geom_point(aes(group = seq_along(final_rank), color = country_fixed), size = 4) +
  geom_line(aes(color = country_fixed), alpha = 0.6, size = 2) +
  scale_x_continuous(breaks = seq(1980, 2016, 4))+
  scale_y_reverse(breaks = seq(1, 10, 1))+ 
  geom_text(aes(label = country_fixed), hjust = -0.5) +
  geom_flag(data = final_df,
            country = final_df$country_code,
            size = 8) +
  labs(x = NULL,
       y = "Rank",
       color = NULL,
       title = "Paralympics medals by country and year") +
  transition_reveal(year) +
  ggthemes::theme_clean() +
  theme(legend.position = "none")

animate(medal_line, fps = 4,
        renderer = gifski_renderer(loop = TRUE)) 
```


```{r}
anim_save("2021_08_03_paralympics.gif")
```

