---
title: "2021_07_20_tidy_tuesday"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(lubridate)
theme_set(theme_light())
```


## Load data

```{r}
drought <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv') %>% 
    mutate(map_date = parse_date(as.character(map_date), format = "%Y%m%d"),
           drought_lvl = fct_recode(drought_lvl,
                                    "No drought" = "None",
                                    "Abnormally dry" = "D0",
                                    "Moderate drought" = "D1",
                                    "Severe drought" = "D2",
                                    "Extreme drought" = "D3",
                                    "Exceptional drought" = "D4"
                                    ),
           drought_lvl = fct_relevel(drought_lvl, c("No drought",
                                                    "Abnormally dry",
                                                    "Moderate drought",
                                                    "Severe drought",
                                                    "Extreme drought",
                                                    "Exceptional drought")))
```

## EDA

```{r}
drought %>% 
    skimr::skim()
```

### Minimum and Maximum recorded date for each state

```{r}
drought %>% 
    group_by(state_abb) %>% 
    summarize(min_date = min(map_date),
              max_date = max(map_date),
              mean_date = mean(map_date),
              median_date = median(map_date)) %>% 
    print(n = Inf)
```

* Each state has the same amount of recorded date

### When was the most prevalent drought (Abnormally dry) in each state?

```{r}
library(geofacet)
drought %>% 
    filter(drought_lvl == "Abnormally dry") %>% 
    group_by(state_abb) %>% 
    slice_max(area_pct, n = 1) %>% 
    ggplot(aes(map_date, area_pct/100)) +
    geom_point() +
    scale_y_continuous(labels = percent) +
    facet_geo(~state_abb) +
    labs(x = "Drought date",
         y = "% of state in drought",
         title = "% of state in drought (abnormally dry) over time") +
    theme(axis.text.x = element_text(angle = 30))
```

* D.C., New Jersey, and Rhode Island have the most prevalent drought (abnormally dry) affecting almost 100% of the state (or special district) on multiple occasions.

### The drought (abnormally dry) hits the most population in each state

```{r}
drought %>% 
    filter(drought_lvl == "Abnormally dry") %>% 
    group_by(state_abb) %>% 
    slice_max(pop_pct, n = 1) %>% 
    ggplot(aes(map_date, pop_pct/100)) +
    geom_point() +
    scale_y_continuous(labels = percent) +
    facet_geo(~state_abb) +
    labs(x = "Drought date",
         y = "% of population in drought",
         title = "% of population in drought (abnormally dry) over time in each state") +
    theme(axis.text.x = element_text(angle = 30))
```

* Similar trend was identified.


### All drought conditions over time

```{r}
drought %>% 
    ggplot(aes(map_date, area_pct/100, fill = drought_lvl)) +
    geom_area() +
    facet_geo(~state_abb, label = "name") +
    scale_fill_viridis_d(option = "A") +
    scale_y_continuous(labels = percent) +
    labs(title = "Droughts in the U.S. (2001 - 2021)",
         subtitle = "% of area in drought",
         fill = "",
         x = "",
         y = "") +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, face = "bold", size = 24),
          plot.subtitle = element_text(hjust = 0.5, face = "bold", size = 16),
          axis.text.x = element_text(angle = 30))
```


### Using plotly to visualize only extreme drought

```{r}
library(plotly)
drought %>%
    filter(drought_lvl == "Extreme drought") %>% 
    group_by(state_abb, year(map_date), month(map_date, label = TRUE)) %>% 
    summarize(`Average area impacted` = mean(area_pct)) %>% 
    ungroup() %>% 
    mutate(year = `year(map_date)`, month = `month(map_date, label = TRUE)`) %>% 
    unite("time", year:month,
                        sep = ", ") %>% 
    plot_ly(type = "choropleth",
            locationmode = "USA-states",
            locations = ~state_abb,
            z = ~`Average area impacted`,
            frame = ~time,
            color = ~`Average area impacted`,
            zauto = FALSE,
            zmin = 0,
            zmax = 100) %>% 
    layout(
    title = "Extreme droughts in the US",
    geo = list(
    scope = 'usa'),
    showlegend = FALSE) %>% 
    colorbar(title = "Avg. % of area impacted")
```


## Visualization at the county level inspired by Katie Press and Andy Baker

```{r}
library(janitor)
library(maps)
library(tibbletime)
library(gganimate)
dr_county <- read_csv("dm_export_20190814_20210727.csv") %>% 
  pivot_longer(cols = None:D4, 
               names_to = "drought_lvl", values_to = "area_pct") %>% 
  clean_names() %>% 
  mutate(drought_lvl = factor(
    case_when(drought_lvl == "None" ~ "No drought",
              drought_lvl == "D0" ~ "Abnormally dry",
              drought_lvl == "D1" ~ "Moderate drought",
              drought_lvl == "D2" ~ "Severe drought",
              drought_lvl == "D3" ~ "Extreme drought",
              drought_lvl == "D4" ~ "Exceptional drought"),
    levels = c("Exceptional drought", 
               "Extreme drought", 
               "Severe drought", 
               "Moderate drought", 
               "Abnormally dry", 
               "No drought"))) 

temp_drought <- dr_county %>% 
  group_by(fips, valid_start) %>% 
  arrange(fips, valid_start, desc(area_pct), drought_lvl) %>% 
  slice(1)

data("county.fips")

county.fips <- county.fips %>% 
  mutate(region = word(polyname, 1, sep = ","),
         subregion = word(polyname, 2, sep = ",")) %>% 
  mutate(subregion = word(subregion, 1, sep = ":")) %>% 
  mutate(fips = str_pad(as.character(fips), side = "left", width = 5, "0"))

map_usa <- map_data("county")

map_usa <- map_usa %>% 
  left_join(county.fips)

series <- tibbletime::create_series("2019-08-13" ~ "2021-07-20", "weekly") %>% 
  mutate(join_col = 1) %>% 
  rename("valid_start" = date)

map_usa <- map_usa %>% 
  mutate(join_col = 1) %>% 
  left_join(series)

map_usa <- map_usa %>% 
  left_join(temp_drought)

temp_nas <- map_usa %>% 
  filter(region == "south dakota", subregion %in% c("shannon", "bennett")) %>% 
  arrange(valid_start, subregion)

temp_nas <- temp_nas %>% 
  fill(c(map_date, state:area_pct)) %>% 
  mutate(county = replace_na(county, "Shannon County")) %>% 
  filter(county == "Shannon County")

map_usa <- map_usa %>% 
  filter(fips != "46113") %>% 
  bind_rows(temp_nas) %>% 
  mutate(valid_start = as.Date(valid_start))


# Mapping -----------------------------------------------------------------

map <- map_usa %>% 
  ungroup() %>% 
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(aes(fill = fct_rev(drought_lvl))) + 
  borders("county", colour = "#FFF8EB", lwd = 0.05) + 
  borders("state", colour = "#FFF8EB", lwd = 1) + 
  coord_map() + 
  scale_fill_manual(values = c("#F1F9F6",
                               "#e9d8a6",
                               "#ee9b00",
                               "#ca6602",
                               "#a61e11",
                               "#641619")) + 
  guides(fill = guide_legend(nrow = 1,
                             label.position = "top",
                             label.hjust = 0.5)) + 
  labs(title = "Drought levels across US counties",
       caption = "Week starting: {current_frame}
       Source: Drought Monitor",
       fill = NULL) + 
  theme_void() + 
  theme(legend.spacing.y = unit(0.35, "cm"),
        legend.spacing.x = unit(3, "cm"),
        legend.position = c(0.5, 0.98),
        legend.key.width = unit(11, "cm"),
        legend.key.height = unit(0.6, "cm"),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(10, 10, 10, 10),
        text = element_text(family = "serif", size = 30),
        plot.caption = element_text(family = "serif", size = 30, hjust = 0.5,
                                    lineheight = 1.2, 
                                    margin = margin(-10, 0, 0, 0)),
        plot.title = element_text(family = "serif", face = "bold", size = 70, 
                                  hjust = 0.5, margin = margin(0, 0, 15, 0)),
        panel.background = element_rect(fill = "#FFF8EB", color = NA),
        plot.background = element_rect(fill = "#FFF8EB", color = NA),
        panel.border = element_blank(),
        plot.margin = margin(50, 10, 30, 10)) + 
  transition_manual(frames = valid_start)

animate(map, fps = 1, width = 2500, height = 1565, 
        renderer = gifski_renderer(loop = TRUE), end_pause = 15)
```


## Similar visulization at state level


```{r}
library(tigris)
states_sf <- states(cb = TRUE)
drought_state <- drought %>% 
  group_by(state_abb, valid_start) %>% 
  arrange(state_abb, valid_start, desc(area_pct), drought_lvl) %>% 
  slice(1) %>% ungroup()
drought_sf <- states_sf %>% 
    right_join(drought_state, by = c("STUSPS" = "state_abb"))
```

```{r}
map <- drought_sf %>%
  filter(valid_start >= "2019-01-01") %>% 
  filter(!STUSPS %in% c("AK", "HI", "PR")) %>% 
  ggplot() +
  geom_sf(aes(fill = drought_lvl)) +
  coord_sf() +
  scale_fill_manual(values = c("#E6E6E6", "#FEF0D9", "#FDCC8A",
                               "#FC8D59", "#E34A33", "#B30000")) +
  guides(fill = guide_legend(nrow = 1,
                             label.position = "top",
                             label.hjust = 0.5)) + 
  labs(title = "Drought level by state 2019-2021",
       caption = "Week: {current_frame}
       Source: Drought Monitor | Viz: tsquall121",
       fill = NULL) + 
  theme_void() +
  theme(legend.spacing.y = unit(0.35, "cm"),
        legend.spacing.x = unit(3, "cm"),
        legend.position = c(0.5, 0.98),
        legend.key.width = unit(11, "cm"),
        legend.key.height = unit(0.6, "cm"),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(10, 10, 10, 10),
        text = element_text(family = "serif", size = 30),
        plot.caption = element_text(family = "serif", size = 30, hjust = 0.5,
                                    lineheight = 1.2, 
                                    margin = margin(-10, 0, 0, 0)),
        plot.title = element_text(family = "serif", face = "bold", size = 70, 
                                  hjust = 0.5, margin = margin(0, 0, 15, 0)),
        panel.border = element_blank(),
        plot.margin = margin(50, 10, 30, 10)) +
  transition_manual(frames = valid_start) 

animate(map, fps = 1, width = 2500, height = 1565, 
        renderer = gifski_renderer(loop = TRUE))

anim_save("2021_07_20_tidy_tuesday_us_drought.gif")
```

